---
title: "R Notebook"
output: html_notebook
---

```{r}
# import relevant libraries
library(readxl)
library(psych)
library(lavaan)
library(semPlot)
library(knitr)
```
```{r}
# Load the data
data <- read_excel("pirls_lesson.xlsx")

head(data, 10)
```


```{r}
# Select the relevant columns for internal and instrumental motivation
internal_motivation <- data[, grep("^enjoy_r[1-6]$", names(data))]
instrumental_motivation <- data[, grep("^reason[1-6]$", names(data))]

# Combine both scales into one data frame
motivation_data <- cbind(internal_motivation, instrumental_motivation)

# Perform factor analysis
# We assume that the items might load onto a single factor (common construct "motivation")
fa_result <- fa(motivation_data, nfactors = 1, rotate = "none")

# Print the factor analysis result
print(fa_result)
```

```{r}
# Check if a single factor solution is sufficient or if multiple factors are needed
fa_parallel <- fa.parallel(motivation_data)
```

The parallel analysis scree plot compares the eigenvalues of the actual data with those from simulated noise to determine the number of meaningful factors or components to retain. The plot indicates that for principal component analysis (PCA), two components should be retained as their eigenvalues are greater than those of the simulated data, while for factor analysis (FA), only one factor is meaningful, with its eigenvalue exceeding the simulated data. Beyond these points, the eigenvalues of the actual data fall below those of the simulated data, suggesting that additional factors or components represent noise rather than meaningful variance. This suggests that the construct being analyzed, such as motivation, may be best represented by a single factor in factor analysis or two components in PCA for dimensionality reduction.


```{r}
# Reverse-score items if they are on a 1-5 scale
reverse_columns <- grep("enjoy_r[1-6]", names(data))
data[, reverse_columns] <- 6 - data[, reverse_columns]

# Check the result
head(data[, reverse_columns])

```

```{r}
# Define the model with two factors (internal and instrumental motivation)
model <- '
  internal =~ enjoy_r1 + enjoy_r2 + enjoy_r3 + enjoy_r4 + enjoy_r5 + enjoy_r6
  instrumental =~ reason1 + reason2 + reason3 + reason4 + reason5 + reason6
'

# Fit the model to the data
fit <- cfa(model, data = data, std.lv = TRUE)

# Display the standardized results
summary(fit, standardized = TRUE, fit.measures = TRUE)

# Extract the standardized correlation between the two factors
correlation <- inspect(fit, "std")$psi[1, 2]
correlation
```

```{r}
# Fit the CFA model to the data
fit <- cfa(model, data = data, std.lv = TRUE)

# Display the fit summary, including standardized estimates and fit measures
summary(fit, standardized = TRUE, fit.measures = TRUE)

# Extract RMSEA and its 90% confidence interval
rmsea <- fitmeasures(fit, c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper"))
rmsea
```

```{r}
# Calculate modification indices
mod_indices <- modindices(fit)

# Filter for potential cross-loadings
cross_loadings <- subset(mod_indices, op == "=~" & !grepl(lhs, rhs))

# Sort by the highest MI
cross_loadings <- cross_loadings[order(-cross_loadings$mi), ]

# Extract the cross-loading with the highest MI
top_cross_loading <- cross_loadings[1, ]

# Display the result
print(top_cross_loading)

```

```{r}
# Get all modification indices
all_mod_indices <- modindices(fit)

# Sort by MI in descending order
sorted_mod_indices <- all_mod_indices[order(-all_mod_indices$mi), ]

# Extract the top two modification indices
top_two_mods <- sorted_mod_indices[1:2, ]

# Display the results
print(top_two_mods)

```

```{r}
# Define the revised CFA model with the two highest modification indices applied
revised_model <- '
  internal =~ enjoy_r1 + enjoy_r2 + enjoy_r3 + enjoy_r4 + enjoy_r5 + enjoy_r6
  instrumental =~ reason1 + reason2 + reason3 + reason4 + reason5 + reason6
  
  # Add cross-loading based on highest MI (example: enjoy_r4 on instrumental)
  instrumental =~ enjoy_r4
  # Add error covariance based on second highest MI (example: enjoy_r2 with enjoy_r3)
  enjoy_r2 ~~ enjoy_r3
'

# Fit the revised CFA model to the data
revised_fit <- cfa(revised_model, data = data, std.lv = TRUE)

# Plot the revised model
semPaths(revised_fit, "std", layout = "tree", residuals = TRUE, 
         intercepts = FALSE, edge.label.cex = 0.8, nCharNodes = 6)

# Extract fit summary to review changes in fit statistics
summary(revised_fit, standardized = TRUE, fit.measures = TRUE)
```

